Bayesian Sparsification of $\cplx$-valued networks

### 2019-11-06 paper/paper.tex
compacted KL-div paragraph

%% drafting macro
\newcommand{\important}[1]{\textbf{\color{red} #1}}
\newcommand{\attn}[2]{\textbf{\color{red} #2~\textsuperscript{\textit{[#1]}}}}
\newcommand{\verify}[1]{\attn{verify}{#1}}
\newcommand{\rewrite}[1]{\attn{rewrite}{#1}}
\newcommand{\todo}[1]{{\color{blue} [TODO]} \important{#1}}

@@

Kullback-Leibler divergence, defined for $q(\omega) \ll p(\omega)$ as
$$
KL(q\| p)
  = \mathbb{E}_{\omega \sim q}
    \log\frac{q(\omega)}{p(\omega)}
  \,, $$
is the natural choice for $\rho$, despite it not being a metric and requiring distributions
with nested support. The main reason is that it satisfies the following identity, which when
everything works, holds for any $q$ and $p$:
\begin{equation} \label{eq:kl-div-master}
  \log p(D)
    % = \mathbb{E}_{q(\omega)} \log \frac{p(D, \omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{q(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    % - \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{\pi(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    - KL(q(\omega) \| \pi(\omega))
    + KL(q(\omega) \| p(\omega \mid D))
    \,.
\end{equation}
This key identity yield an equivalent problem for \eqref{eq:variational-progam}:
maximizing the \textit{evidence lower bound} (ELBO)
\begin{equation} \label{eq:elbo_general}
\begin{aligned}
  & \underset{\theta, \phi, \lambda}{\text{maximize}}
    & \mathcal{L}(\theta; \phi, \lambda)
      & = \mathbb{E}_{\omega \sim q_{\theta}}
          \log p_{\phi}(D \mid \omega)
        - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
        % \mathbb{E}_{\omega \sim q_{\theta}}
          % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % \\
  % & &
  %     & = \mathbb{E}_{\omega \sim q_{\theta}}
  %         \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %       + \mathbb{H}(q_{\theta})
  \,,
\end{aligned}
\end{equation}
where the prior and likelihood are allowed to depend on explicit parameters. The most
commonly used algorithm to optimize \eqref{eq:elbo_general} is to use EM-algorithm.
Other variational objectives are possible, provided $p(\omega \mid D)$ is probed only
through $\log p(D \mid \omega)$ and $\log \pi(\omega)$, \cite{ranganath_operator_2018}.

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \cite{kingma_auto-encoding_2014},
is an differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which makes
it possible to employ stochastic gradient-based optimization methods. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation} \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid \omega_{il})
          \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta}(\omega) \| \pi(\omega))
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be also
replaced by differentiable unbiased Monte-Carlo estimators, \cite{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation} \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.

>>>>

%% drafting macro
\newcommand{\important}[1]{\textbf{\colorbox{red}{#1}}}
\newcommand{\attn}[2]{\textbf{\color{red} #2~\textsuperscript{\textit{[#1]}}}}
\newcommand{\verify}[1]{\attn{verify}{#1}}
\newcommand{\rewrite}[1]{\attn{rewrite}{#1}}
\newcommand{\todo}[1]{{\color{blue} [TODO]} \important{#1}}

@@

Kullback-Leibler divergence of density $q$ from $p$ is
\begin{equation} \label{eq:kl-div-def}
  KL(q \| p)
    % = \int \frac{dQ}{dP} \log{\frac{dQ}{dP}} P(d\omega)
    % = \int \frac{q(\omega)}{p(\omega)} \log{\frac{q(\omega)}{p(\omega)}} p(\omega) d\omega
    % = \mathbb{E}_{\omega \sim q}
    %   \log\frac{q(\omega)}{p(\omega)}
    = \mathbb{E}_{\omega \sim q}
      \log{q(\omega)}
    - \mathbb{E}_{\omega \sim q}
      \log{p(\omega)}
    \,,
\end{equation}
% and, despite it not being a metric and requiring distributions with nested support,
and is the natural choice for $\rho$, the main reason being that it satisfies the following
identity for well-behaved $q$ and $p$:
\begin{equation} \label{eq:kl-div-master}
  \log p(D)
    % = \mathbb{E}_{q(\omega)} \log \frac{p(D, \omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{q(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    % - \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{\pi(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    - KL(q(\omega) \| \pi(\omega))
    + KL(q(\omega) \| p(\omega \mid D))
    \,.
\end{equation}
This key identity yield an equivalent problem for \eqref{eq:variational-progam}:
maximizing the \textit{evidence lower bound} (ELBO)
\begin{equation} \label{eq:elbo_general}
\begin{aligned}
  & \underset{\theta, \phi, \lambda}{\text{maximize}}
    & \mathcal{L}(\theta; \phi, \lambda)
      & = \mathbb{E}_{\omega \sim q_{\theta}}
          \log p_{\phi}(D \mid \omega)
        - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
        % \mathbb{E}_{\omega \sim q_{\theta}}
          % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % \\
  % & &
  %     & = \mathbb{E}_{\omega \sim q_{\theta}}
  %         \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %       + \mathbb{H}(q_{\theta})
  \,,
\end{aligned}
\end{equation}
where the prior and likelihood are allowed to depend on explicit parameters. The most
commonly used algorithm to optimize \eqref{eq:elbo_general} is to use EM-algorithm.
Other variational objectives are possible, provided $p(\omega \mid D)$ is probed only
through $\log p(D \mid \omega)$ and $\log \pi(\omega)$, \cite{ranganath_operator_2018}.

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \cite{kingma_auto-encoding_2014},
is an differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which makes
it possible to employ stochastic gradient-based optimization methods. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation} \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid \omega_{il})
          \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta}(\omega) \| \pi(\omega))
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be also
replaced by differentiable unbiased Monte-Carlo estimators, \cite{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation} \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.
%
\todo{DSVI}
